\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-08 tex/paper.tex
switch to nicer APS template

\documentclass{article}

\PassOptionsToPackage{numbers, compress}{natbib}

\usepackage[final]{nips_2016}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\title{Adversarial Training against Systematic Uncertainty}

\author{
  Gilles Louppe \\
  New York University\\
  \texttt{g.louppe@nyu.edu} \\
  \And
  Michael Kagan\\
  SLAC National Accelerator Laboratory\\
  \texttt{makagan@slac.stanford.edu}\\
  \And
  Kyle Cranmer \\
  New York University\\
  \texttt{kyle.cranmer@nyu.edu}
}

@@

@@

Lorem ipsum dolor sit amet, consectetur adipiscing elit...

@@

Since we do not have training tuples $(X_{z}(\omega),
X_{z^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{equation}\label{eqn:criterion-measure}
    P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\}) \text{ for all $y \in {\cal Y}$}.
\end{equation}
In words, we are looking for a predictive function $f$ which is a pivotal
quantity \citep{degroot1986probability} with respect to the nuisance parameter.
That is, such that  the distribution of $f(X_z; \theta_f)$ is invariant
with respect to the value $z$ of the nuisance. Note that a function $f$ for which
Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true, since the
sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
for the equality to hold.
In order to simplify notations,
and as only Eqn.~\ref{eqn:criterion-measure} is
of direct interest in this work, we denote from here on
the pivotal quantity criterion as
\begin{equation}\label{eqn:criterion}
    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime ) \text{ for all $z,z^\prime \in  {\cal Z}$.}
\end{equation}

@@

Adversarial training was first proposed by \cite{goodfellow2014generative} as a
way to build a generative model capable of producing samples from random noise
$z \sim p_Z$. More specifically, the authors pit a generative model $g: \mathbb{R}
\mapsto \mathbb{R}^p$ against an adversary classifier $d : \mathbb{R}^p \mapsto \{
0, 1\}$ whose antagonistic objective is to recognize real data $X$ from generated data $g(Z)$. Both
models $g$ and $d$ are trained simultaneously, in such a way that $g$ learns to
produce samples that are difficult to identify by $d$, while $d$ incrementally
adapts to changes in $g$. At the equilibrium, $g$ models a distribution whose
samples can be identified by $d$ only by chance. That is, assuming enough
capacity in $d$ and  $g$, the distribution $p_{g(Z)}$ eventually converges
towards the real distribution $p_X$.

@@

\begin{algorithm}[t]
\caption{Adversarial training of a classifier $f$ against an adversary $r$.\\
{\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$\\
{\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$\\
{\it Hyper-parameters:} Number $T$ of training iterations, Number $K$ of gradient steps to update $r$.}
\label{alg:adversarial-training}
\begin{algorithmic}[1]
    \For{$t=1$ to $T$}
        \For{$k=1$ to $K$} \Comment{Update $r$}
            \State{Sample minibatch $\{x_m, z_m \}_{m=1}^M$} of size $M$;
            \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
            $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|f(x_m;\theta_f))  ;$$}
        \EndFor
        \State{Sample minibatch $\{x_m, y_m, z_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
        \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
        $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|f(x_m;\theta_f))  \right],$$
        \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-f(x_m;\theta_f)) + 1(y_m=1)f(x_m;\theta_f))$;}
    \EndFor
\end{algorithmic}
\end{algorithm}

@@

\begin{proposition}\label{prop:1}
Let $\theta_f$ be fixed and $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
\theta_r)$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
\end{proposition}

\begin{proof}
The optimal parameters $\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
\theta_r)$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
By assumption, $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$,
and therefore $p(z|f(X;\theta_f)) = p(z)$.
Using the Bayes' rule, we write
\begin{align*}
    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)} = p(f(X;\theta_f)),
\end{align*}
which holds for all $z \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
\end{proof}

@@

\begin{proof}
For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal L}_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the expected negative log-likelihood ${\cal L}_f$ of $Y|X$
    reduces to its minimum value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
   Z}$ since we require a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
   over continuous distributions on the same bounded support is also realized by the uniform distribution over that support.
\end{itemize}
Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}

@@

\glnote{Alternative to learning a parameterized classifier for LLR.}

\glnote{Similar to domain adaptation, but with infinitely many domains,
as parameterized by $Z$, also related to transfer learning.}

\glnote{Other applications: removing implicit bias in data (e.g. gender bias).}

@@

\section*{Acknowledgments}

\bibliographystyle{ieeetr}
\bibliography{bibliography.bib}

>>>>

\documentclass[twocolumn,preprintnumbers,superscriptaddress,nofootinbib,aps,floatfix,letterpaper]{revtex4-1}

\usepackage[utf8]{inputenc}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\usepackage{bm}
\usepackage{cancel}
\usepackage{bbold}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{color}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

@@

\title{\Large{Adversarial Training against Systematic Uncertainty}}
\vspace{1cm}
\author{\small{\bf Gilles Louppe}\thanks{\texttt{g.louppe@nyu.edu}}}
\affiliation{New York University}
\author{\small{\bf Michael Kagan}\thanks{\texttt{makagan@slac.stanford.edu}}}
\affiliation{SLAC National Accelerator Laboratory}
\author{\small{\bf Kyle Cranmer}\thanks{\texttt{kyle.cranmer@nyu.edu}}}
\affiliation{New York University}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam commodo, enim
vitae facilisis pretium, ligula justo aliquet lectus, non interdum erat lorem et
nisi. Donec pharetra lectus in magna pellentesque vehicula. Praesent dapibus
lorem sed enim lacinia, a vestibulum sapien mattis. Sed ornare mollis aliquet.
Nulla tempor lacinia tortor, in rhoncus augue porta nec. Morbi sed convallis
nibh, eu hendrerit turpis. Curabitur sit amet rhoncus purus. Curabitur eget
magna lorem. Phasellus pretium nisi quis est tincidunt, faucibus vulputate augue
viverra. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Proin a urna a
ex egestas pulvinar. Curabitur facilisis a dui nec fringilla. Duis bibendum
fermentum quam. Donec eget urna eget mi placerat malesuada ut et lacus. Nulla ac
risus scelerisque, congue eros vitae, dignissim risus. Fusce purus ante,
vulputate sed magna et, varius imperdiet leo.

@@

@@

Since we do not have training tuples $(X_{z}(\omega),
X_{z^\prime}(\omega))$ (for the same unknown $\omega$), we propose instead to
solve the closely related problem of finding a predictive function $f$ such that
\begin{align}\label{eqn:criterion-measure}
    &P(\{ \omega | f(X_{z}(\omega) ; \theta_f) = y \} ) \nonumber \\
    &\indent = P( \{ \omega' | f(X_{z^\prime}(\omega') ; \theta_f) = y\})
\end{align}
for all $y \in {\cal Y}$. In words, we are looking for a predictive function $f$
which is a pivotal quantity \citep{degroot1986probability} with respect to the
nuisance parameter. That is, such that  the distribution of $f(X_z; \theta_f)$
is invariant with respect to the value $z$ of the nuisance. Note that a function
$f$ for which Eqn.~\ref{eqn:criterion-true} is true necessarily satisfies
Eqn.~\ref{eqn:criterion-measure}. In general, the converse is however not true,
since the sets of samples $\{ \omega | f(X_{z}(\omega); \theta_f) = y \}$ and $\{
\omega' | f(X_{z^\prime}(\omega'); \theta_f) = y \}$ do not need to be the same
for the equality to hold. In order to simplify notations, and as only
Eqn.~\ref{eqn:criterion-measure} is of direct interest in this work, we denote
from here on the pivotal quantity criterion as
\begin{equation}\label{eqn:criterion}
    p(f(X ; \theta_f) | z ) = p(f(X ; \theta_f) | z^\prime )
\end{equation}
for all $z,z^\prime \in  {\cal Z}$.

@@

Adversarial training was first proposed by \citep{goodfellow2014generative} as a
way to build a generative model capable of producing samples from random noise
$z \sim p_Z$. More specifically, the authors pit a generative model $g:
\mathbb{R} \mapsto \mathbb{R}^p$ against an adversary classifier $d :
\mathbb{R}^p \mapsto \{ 0, 1\}$ whose antagonistic objective is to recognize
real data $X$ from generated data $g(Z)$. Both models $g$ and $d$ are trained
simultaneously, in such a way that $g$ learns to produce samples that are
difficult to identify by $d$, while $d$ incrementally adapts to changes in $g$.
At the equilibrium, $g$ models a distribution whose samples can be identified by
$d$ only by chance. That is, assuming enough capacity in $d$ and  $g$, the
distribution $p_{g(Z)}$ eventually converges towards the real distribution
$p_X$.

@@

\begin{figure*}
    \begin{minipage}{\linewidth}
    \begin{algorithm}[H]
    \caption{Adversarial training of a classifier $f$ against an adversary $r$.}

    \begin{flushleft}
        {\it Inputs:} training data $\{ x_i, y_i, z_i \}_{i=1}^N$;\\
        {\it Outputs:} $\smash{\hat\theta_f}, \smash{\hat\theta_r}$;\\
        {\it Hyper-parameters:} Number $T$ of training iterations,
                                Number $K$ of gradient steps to update $r$.
    \end{flushleft}

    \label{alg:adversarial-training}
    \begin{algorithmic}[1]
        \For{$t=1$ to $T$}
            \For{$k=1$ to $K$} \Comment{Update $r$}
                \State{Sample minibatch $\{x_m, z_m \}_{m=1}^M$} of size $M$;
                \State{With $\theta_f$ fixed, update $r$ by ascending its stochastic gradient $\nabla_{\theta_r} E(\theta_f, \theta_r) :=$
                $$\nabla_{\theta_r} \sum_{m=1}^M \log p_{\theta_r}(z_m|f(x_m;\theta_f))  ;$$}
            \EndFor
            \State{Sample minibatch $\{x_m, y_m, z_m \}_{m=1}^M$} of size $M$; \Comment{Update $f$}
            \State{With $\theta_r$ fixed, update $f$ by descending its stochastic gradient $\nabla_{\theta_f} E(\theta_f, \theta_r) :=$
            $$\nabla_{\theta_f}  \sum_{m=1}^M \left[ -\log p_{\theta_f}(y_m|x_m)  +\log p_{\theta_r}(z_m|f(x_m;\theta_f))  \right],$$
            \indent where $p_{\theta_f}(y_m|x_m)$ denotes $1(y_m=0)(1-f(x_m;\theta_f)) + 1(y_m=1)f(x_m;\theta_f))$;}
        \EndFor
    \end{algorithmic}
    \end{algorithm}
    \end{minipage}
\end{figure*}

@@

\begin{proposition}\label{prop:1}
Let $\hat\theta_r = \arg \max_{\theta_r} E(\theta_f,
\theta_r)$ for fixed $\theta_f$. If $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$
for all $z \in {\cal Z}$, then $f$ is a pivotal quantity.
\end{proposition}

\begin{proof}
The optimal parameters $$\smash{\hat\theta}_r =
\arg \max_{\theta_r} E(\theta_f, \theta_r) = \arg \min_{\theta_r} {\cal L}_r(\theta_f,
\theta_r)$$ are such that $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$.
By assumption, $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$,
and therefore $p(z|f(X;\theta_f)) = p(z)$.
Using the Bayes' rule, we write
\begin{align*}
    p(f(X;\theta_f)|z) &= \frac{ p(z|f(X;\theta_f)) p(f(X;\theta_f)) } { p(z)}\\
                       &= p(f(X;\theta_f)),
\end{align*}
which holds for all $z \in {\cal Z}$ and implies that $f$ is a pivotal quantity.
\end{proof}

@@

\begin{proof}

For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} {\cal
L}_r(\theta_f, \theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) =
p(z|f(X;\theta_f))$ and ${\cal L}_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = {\cal
L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower
bound $$H(p_{Y|X}) - H(p_Z) \leq {\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$$
where the equality holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f}
E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the expected
    negative log-likelihood ${\cal L}_f$ of $Y|X$ reduces to its minimum
    value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
   Z}$ since we require a uniform prior by construction. Note that in the
   continuous case, the supremum of the differential entropy
   over continuous distributions on the same bounded support is also realized
   by the uniform distribution over that support.
\end{itemize}
Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) =
p(z)$ and therefore that at this point, because of Proposition~\ref{prop:1}, the
optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.

\end{proof}

@@

\glnote{Alternative to learning a parameterized classifier for LLR.}
\glnote{Similar to domain adaptation, but with infinitely many domains,
as parameterized by $Z$, also related to transfer learning.}
\glnote{Other applications: removing implicit bias in data (e.g. gender bias).}

@@

\begin{acknowledgments}
The authors would like to thank...
\end{acknowledgments}

\bibliography{bibliography.bib}
